---
title: "Analyse de la Fréquence Allélique et Déduction du Pourcentage Tumoral"
author: "Léo Zwilling"
date: "`r Sys.Date()`"
output: pdf_document
---

## Introduction

This document performs an analysis of allele frequency and estimates the tumoral percentage of the sample. It includes data preparation, visualization, and summary statistics.

## Data Visualization with GT

We will first create a GT table to visualize the data with conditional color formatting based on the LOH classification.

```{r gt_table, echo=FALSE}
library(dplyr)
library(gt)
library(here)
library(ggplot2)

# Load the cleaned data
cons_tum <- readRDS(file = here("data/cons_tum_cleaned.rds"))

# Create a gt table with color formatting
# Apply conditional color formatting row by row
cons_tum %>%
  gt() %>%
  tab_style(
    style = cell_fill(color = "#B22222"),  # Darker red
    locations = cells_body(
      columns = vars(LOH, Coverage.tum, `%tumoral`),
      rows = LOH == "CIS"
    )
  ) %>%
  tab_style(
    style = cell_fill(color = "#228B22"),  # Darker green
    locations = cells_body(
      columns = vars(LOH, Coverage.tum, `%tumoral`),
      rows = LOH == "TRANS"
    )
  )

``` 

## Data Preparation

We extract the %tumoral values based on LOH classification and calculate the mean and standard deviation for each classification.

```{r data_preparation, echo=FALSE} 
# Extract lists of %tumoral values based on LOH classification
data_for_plot <- cons_tum %>%
  filter(!is.na(`%tumoral`)) %>%
  select(LOH, `%tumoral`)

# Calculate mean and standard deviation for each LOH classification
summary_stats <- data_for_plot %>%
  group_by(LOH) %>%
  summarise(
    Mean = mean(`%tumoral`, na.rm = TRUE),
    SD = sd(`%tumoral`, na.rm = TRUE)
  )

# Calculate mean of the two classes
mean_CIS_TRANS <- cons_tum %>%
  filter(LOH %in% c("CIS", "TRANS")) %>%
  summarise(Mean = mean(`%tumoral`, na.rm = TRUE)) %>%
  pull(Mean)

``` 

## Boxplot Visualization

We create a boxplot of %tumoral by LOH classification, displaying the mean and standard deviation.


```{r boxplot, echo=FALSE} 

# Create the boxplot
ggplot(data_for_plot, aes(x = LOH, y = `%tumoral`, fill = LOH)) +
  geom_boxplot(varwidth = TRUE, fill="plum") +
  geom_point(data = summary_stats, aes(x = LOH, y = Mean), color = "#4D4D4D", size = 3, shape = 20, show.legend = FALSE) +
  geom_text(data = summary_stats, aes(x = LOH, y = Mean, label = paste("Mean:", round(Mean, 2), "±", round(SD, 2))), 
            vjust = -0.5, hjust = 1, color = "#4D4D4D", size = 3.5) +  
  labs(title = "Box Plot",
       subtitle = "% Tumoral by LOH Classification",
       x = NULL,
       y = "% Tumoral") +
  theme_minimal()

```

## Results

The calculated mean for the CIS and TRANS classes combined is:


```{r result, echo=FALSE} 
mean_CIS_TRANS
``` 








